Modern BTree\\ techniques

### 2020-01-14 btree/presentation.tex
Partitioned btree

\tikzset{rum-node/.style={draw,circle,fill=white,minimum width=2cm}}
\tikzset{rum-extra-node/.style={rectangle, draw, dashed, rounded corners, fill=white }}
\tikzset{btree-key/.style={minimum height=1cm, pattern=north west lines, draw}}
\tikzset{btree-pointer/.style={minimum height=1cm, draw}}
\tikzset{btree-hide/.style={draw opacity=0, line width=0, pattern=none}}
\tikzset{btree-empty/.style={btree-pointer, dashed}}
\tikzset{btree-leaf/.style={fill=greenGood}}
\tikzset{btree-branch/.style={fill=gray!20}}
\tikzset{btree-line/.style={line width=0.5pt}}
\tikzset{btree-path/.style={btree-pointer, minimum width=0.8cm, fill=red, opacity=0.5}}
\tikzset{>=latex}
\tikzset{btree-high-key/.style={btree-key, pattern=north west lines, pattern color=red, draw=red}}
\tikzset{btree-key-prefix/.style={btree-key, minimum height=0.2cm, fill=red, opacity=0.5}}
\tikzset{btree-key-compare/.style={btree-key, minimum height=0.6cm, fill=blue, opacity=0.5}}
\tikzset{arrow-pointer/.style={
        single arrow,
        minimum height=1.5cm,
        inner sep=3pt,
        line width=1pt,
        draw,
        color=gray,
        single arrow tip angle=45,
        single arrow head extend=0.1cm
    }
}
\tikzset{btree-indirection/.style={btree-key, minimum width=0.2cm, inner sep=0, pattern=none}}
\tikzset{btree-page-header/.style={btree-key, minimum width=0.5cm, inner sep=0, pattern=none}}
\tikzset{cpu-cache/.style={draw, minimum width=1cm, minimum height=1cm, text width=1cm, align=center, rounded corners}}
\tikzset{cpu-cache-hide/.style={cpu-cache, draw=none, color=white}}

@@

            \coordinate[below=1cm of key1] (start);
            \draw[->, btree-line] (start.south)
                .. controls ([xshift=1cm] start) and ([yshift=-1cm] key6) ..
                (key6.south);
            \draw[->, btree-line] (key6.south)
                .. controls ([yshift=-1cm] key6) and ([yshift=-1cm] key9) ..
                (key9.south);
            \draw[->, btree-line] (key9.south)
                .. controls ([yshift=-0.9cm] key9) and ([yshift=-0.9cm] key10) ..
                (key10.south);

@@

            \coordinate[below=1cm of key6] (start);
            \draw[->, btree-line] (start.south)
                .. controls ([xshift=1cm] start) and ([yshift=-1cm] key10) ..
                (key10.south);

@@

>>>>

\tikzset{rum-node/.style={draw,circle,fill=white,minimum width=2cm}}
\tikzset{rum-extra-node/.style={rectangle, draw, dashed, rounded corners, fill=white }}
\tikzset{btree-key/.style={minimum height=1cm, pattern=north west lines, draw}}
\tikzset{btree-pointer/.style={minimum height=1cm, draw}}
\tikzset{btree-hide/.style={draw opacity=0, line width=0, pattern=none}}
\tikzset{btree-empty/.style={btree-pointer, dashed}}
\tikzset{btree-leaf/.style={fill=greenGood}}
\tikzset{btree-branch/.style={fill=gray!20}}
\tikzset{btree-line/.style={line width=0.5pt}}
\tikzset{btree-path/.style={btree-pointer, minimum width=0.8cm, fill=red, opacity=0.5}}
\tikzset{>=latex}
\tikzset{btree-high-key/.style={btree-key, pattern=north west lines, pattern color=red, draw=red}}
\tikzset{btree-key-prefix/.style={btree-key, minimum height=0.2cm, fill=red, opacity=0.5}}
\tikzset{btree-key-compare/.style={btree-key, minimum height=0.6cm, fill=blue, opacity=0.5}}
\tikzset{arrow-pointer/.style={
        single arrow,
        minimum height=1.5cm,
        inner sep=3pt,
        line width=1pt,
        draw,
        color=gray,
        single arrow tip angle=45,
        single arrow head extend=0.1cm
    }
}
\tikzset{btree-indirection/.style={btree-key, minimum width=0.2cm, inner sep=0, pattern=none}}
\tikzset{btree-page-header/.style={btree-key, minimum width=0.5cm, inner sep=0, pattern=none}}
\tikzset{cpu-cache/.style={draw, minimum width=1cm, minimum height=1cm, text width=1cm, align=center, rounded corners}}
\tikzset{cpu-cache-hide/.style={cpu-cache, draw=none, color=white}}
\tikzset{btree-node/.style={draw, inner sep=0, minimum width=1cm, minimum height=1cm, rounded corners}}
\tikzset{btree-partition1/.style={draw, inner sep=0, minimum width=1cm, minimum height=1cm, rounded corners, fill=red!20}}
\tikzset{btree-partition2/.style={draw, inner sep=0, minimum width=1cm, minimum height=1cm, rounded corners, fill=green!20}}
\tikzset{btree-partition3/.style={draw, inner sep=0, minimum width=1cm, minimum height=1cm, rounded corners, fill=blue!20}}

@@

            \coordinate[below=2cm of key1] (start);
            \draw[->, btree-line] (start.south)
                .. controls ([xshift=1cm] start) and ([yshift=-3cm] key6) ..
                (key6.south);
            \draw[->, btree-line] (key6.south)
                .. controls ([yshift=-3cm] key6) and ([yshift=-3cm] key9) ..
                (key9.south);
            \draw[->, btree-line] (key9.south)
                .. controls ([yshift=-2cm] key9) and ([yshift=-2cm] key10) ..
                (key10.south);

@@

            \coordinate[below=2cm of key6] (start);
            \draw[->, btree-line] (start.south)
                .. controls ([xshift=1cm] start) and ([yshift=-3cm] key10) ..
                (key10.south);

@@

    \begin{center}
        \includegraphics[width=0.65\textwidth,center]{waves.png}
    \end{center}
\end{frame}

    \begin{center}
    \textbf{Partitioned B-tree}
    \vspace{1cm}

    \begin{overprint}[8cm]
        \onslide<1>
        \begin{tikzpicture}[]
            \node[btree-node] (node1) {};

            \node[btree-node, below=0.5cm of node1.south west, xshift=-2.5cm] (node2) {};
            \node[btree-node, below=0.5cm of node1.south] (node3) {};
            \node[btree-node, below=0.5cm of node1.south east, xshift=2.5cm] (node4) {};

            \node[btree-node, below=0.5cm of node2.south west, xshift=-0.2cm] (node5) {};
            \node[btree-node, below=0.5cm of node2.south east, xshift=0.2cm] (node6) {};

            \node[btree-node, below=0.5cm of node3.south west, xshift=-0.2cm] (node7) {};
            \node[btree-node, below=0.5cm of node3.south east, xshift=0.2cm] (node8) {};

            \node[btree-node, below=0.5cm of node4.south west, xshift=-0.2cm] (node9) {};
            \node[btree-node, below=0.5cm of node4.south east, xshift=0.2cm] (node10) {};

            \draw[->, btree-line] (node1) -- (node2);
            \draw[->, btree-line] (node1) -- (node3);
            \draw[->, btree-line] (node1) -- (node4);

            \draw[->, btree-line] (node2) -- (node5);
            \draw[->, btree-line] (node2) -- (node6);

            \draw[->, btree-line] (node3) -- (node7);
            \draw[->, btree-line] (node3) -- (node8);

            \draw[->, btree-line] (node4) -- (node9);
            \draw[->, btree-line] (node4) -- (node10);
        \end{tikzpicture}

        \onslide<2>
        \begin{tikzpicture}[]
            \node[btree-node] (node1) {};

            \node[btree-node, below=0.5cm of node1.south west, xshift=-2.5cm] (node2) {};
            \node[btree-node, below=0.5cm of node1.south] (node3) {};
            \node[btree-partition1, below=0.5cm of node1.south east, xshift=2.5cm] (node4) {};

            \node[btree-node, below=0.5cm of node2.south west, xshift=-0.2cm] (node5) {};
            \node[btree-node, below=0.5cm of node2.south east, xshift=0.2cm] (node6) {};

            \node[btree-node, below=0.5cm of node3.south west, xshift=-0.2cm] (node7) {};
            \node[btree-node, below=0.5cm of node3.south east, xshift=0.2cm] (node8) {};

            \node[btree-partition1, below=0.5cm of node4.south west, xshift=-0.2cm] (node9) {};
            \node[btree-partition1, below=0.5cm of node4.south east, xshift=0.2cm] (node10) {};

            \draw[->, btree-line] (node1) -- (node2);
            \draw[->, btree-line] (node1) -- (node3);
            \draw[->, btree-line] (node1) -- (node4);

            \draw[->, btree-line] (node2) -- (node5);
            \draw[->, btree-line] (node2) -- (node6);

            \draw[->, btree-line] (node3) -- (node7);
            \draw[->, btree-line] (node3) -- (node8);

            \draw[->, btree-line] (node4) -- (node9);
            \draw[->, btree-line] (node4) -- (node10);
        \end{tikzpicture}

        \onslide<3>
        \begin{tikzpicture}[]
            \node[btree-node] (node1) {};

            \node[btree-partition1, below=0.5cm of node1.south west, xshift=-2.5cm] (node2) {};
            \node[btree-partition2, below=0.5cm of node1.south] (node3) {};
            \node[btree-partition3, below=0.5cm of node1.south east, xshift=2.5cm] (node4) {};

            \node[btree-partition1, below=0.5cm of node2.south west, xshift=-0.2cm] (node5) {};
            \node[btree-partition1, below=0.5cm of node2.south east, xshift=0.2cm] (node6) {};

            \node[btree-partition2, below=0.5cm of node3.south west, xshift=-0.2cm] (node7) {};
            \node[btree-partition2, below=0.5cm of node3.south east, xshift=0.2cm] (node8) {};

            \node[btree-partition3, below=0.5cm of node4.south west, xshift=-0.2cm] (node9) {};
            \node[btree-partition3, below=0.5cm of node4.south east, xshift=0.2cm] (node10) {};

            \draw[->, btree-line] (node1) -- (node2);
            \draw[->, btree-line] (node1) -- (node3);
            \draw[->, btree-line] (node1) -- (node4);

            \draw[->, btree-line] (node2) -- (node5);
            \draw[->, btree-line] (node2) -- (node6);

            \draw[->, btree-line] (node3) -- (node7);
            \draw[->, btree-line] (node3) -- (node8);

            \draw[->, btree-line] (node4) -- (node9);
            \draw[->, btree-line] (node4) -- (node10);

            \node[arrow-pointer, below=3cm of node2.center, rotate=90, yshift=0.2cm] {};
            \node[arrow-pointer, below=3cm of node3.center, rotate=90, yshift=0.2cm] {};
            \node[arrow-pointer, below=3cm of node4.center, rotate=90, yshift=0.2cm] {};
            \node[below=3cm of node3.center, minimum height=1cm] {};

        \end{tikzpicture}
        \linespread{0.5}
        \color{black}\fontsize{6pt}{0}\selectfont
            G. Graefe, "Sorting and indexing with partitioned B-Trees",
            Classless Inter Domain Routing, 2003
        \linespread{1.5}

        \onslide<4>
        \begin{tikzpicture}[]
            \node[btree-node] (node1) {};

            \node[btree-partition1, below=0.5cm of node1.south west, xshift=-2.5cm] (node2) {};
            \node[btree-partition2, below=0.5cm of node1.south] (node3) {};
            \node[btree-partition3, below=0.5cm of node1.south east, xshift=2.5cm] (node4) {};

            \node[btree-partition1, below=0.5cm of node2.south west, xshift=-0.2cm] (node5) {};
            \node[btree-partition1, below=0.5cm of node2.south east, xshift=0.2cm] (node6) {};

            \node[btree-partition2, below=0.5cm of node3.south west, xshift=-0.2cm] (node7) {};
            \node[btree-partition2, below=0.5cm of node3.south east, xshift=0.2cm] (node8) {};

            \node[btree-partition3, below=0.5cm of node4.south west, xshift=-0.2cm] (node9) {};
            \node[btree-partition3, below=0.5cm of node4.south east, xshift=0.2cm] (node10) {};

            \draw[->, btree-line] (node1) -- (node2);
            \draw[->, btree-line] (node1) -- (node3);
            \draw[->, btree-line] (node1) -- (node4);

            \draw[->, btree-line] (node2) -- (node5);
            \draw[->, btree-line] (node2) -- (node6);

            \draw[->, btree-line] (node3) -- (node7);
            \draw[->, btree-line] (node3) -- (node8);

            \draw[->, btree-line] (node4) -- (node9);
            \draw[->, btree-line] (node4) -- (node10);

            \draw[->, btree-line] ([yshift=-2cm]node2.center)
                .. controls ([yshift=-3cm] node2) and ([yshift=-3cm] node3) ..
                ([yshift=-2cm]node3.center);
            \draw[->, btree-line] ([yshift=-2cm]node4.center)
                .. controls ([yshift=-3cm] node4) and ([yshift=-3cm] node3) ..
                ([yshift=-2cm]node3.center);
            \node[below=3cm of node3.center, minimum height=1cm] {};

        \end{tikzpicture}
        \linespread{0.5}
        \color{black}\fontsize{6pt}{0}\selectfont
            G. Graefe, "Sorting and indexing with partitioned B-Trees",
            Classless Inter Domain Routing, 2003
        \linespread{1.5}

        \onslide<5>
        \begin{tikzpicture}[]
            \node[btree-node] (node1) {};

            \node[btree-node, below=0.5cm of node1.south west, xshift=-2.5cm] (node2) {};
            \node[btree-node, below=0.5cm of node1.south] (node3) {};
            \node[btree-partition1, below=0.5cm of node1.south east, xshift=2.5cm] (node4) {};

            \node[btree-node, below=0.5cm of node2.south west, xshift=-0.2cm] (node5) {};
            \node[btree-node, below=0.5cm of node2.south east, xshift=0.2cm] (node6) {};

            \node[btree-node, below=0.5cm of node3.south west, xshift=-0.2cm] (node7) {};
            \node[btree-node, below=0.5cm of node3.south east, xshift=0.2cm] (node8) {};

            \node[btree-partition1, below=0.5cm of node4.south west, xshift=-0.2cm] (node9) {};
            \node[btree-partition1, below=0.5cm of node4.south east, xshift=0.2cm] (node10) {};

            \draw[->, btree-line] (node1) -- (node2);
            \draw[->, btree-line] (node1) -- (node3);
            \draw[->, btree-line] (node1) -- (node4);

            \draw[->, btree-line] (node2) -- (node5);
            \draw[->, btree-line] (node2) -- (node6);

            \draw[->, btree-line] (node3) -- (node7);
            \draw[->, btree-line] (node3) -- (node8);

            \draw[->, btree-line] (node4) -- (node9);
            \draw[->, btree-line] (node4) -- (node10);

            \node[arrow-pointer, right=2cm of node9.center, rotate=-135, xshift=-2.7cm, yshift=-1cm] {};
            \node[draw, behind path, fit=(node4)(node9)(node10), rounded corners, dashed] {};
            \node[below=3cm of node3.center, minimum height=1cm] {};

        \end{tikzpicture}
        \linespread{0.5}
        \color{black}\fontsize{6pt}{0}\selectfont
            Riegger, Christian Vincon, Tobias Petrov, Ilia. (2017).
            Write-optimized indexing with partitioned b-trees. 296-300.
            10.1145/3151759.3151814.
        \linespread{1.5}

        \onslide<6>
        \begin{tikzpicture}[]
            \node[btree-node] (node1) {};

            \node[btree-node, below=0.5cm of node1.south west, xshift=-2.5cm] (node2) {};
            \node[btree-node, below=0.5cm of node1.south] (node3) {};
            \node[btree-partition1, below=0.5cm of node1.south east, xshift=2.5cm] (node4) {};

            \node[btree-node, below=0.5cm of node2.south west, xshift=-0.2cm] (node5) {};
            \node[btree-node, below=0.5cm of node2.south east, xshift=0.2cm] (node6) {};

            \node[btree-node, below=0.5cm of node3.south west, xshift=-0.2cm] (node7) {};
            \node[btree-node, below=0.5cm of node3.south east, xshift=0.2cm] (node8) {};

            \node[btree-partition1, below=0.5cm of node4.south west, xshift=-0.2cm] (node9) {};
            \node[btree-partition1, below=0.5cm of node4.south east, xshift=0.2cm] (node10) {};

            \draw[->, btree-line] (node1) -- (node2);
            \draw[->, btree-line] (node1) -- (node3);
            \draw[->, btree-line] (node1) -- (node4);

            \draw[->, btree-line] (node2) -- (node5);
            \draw[->, btree-line] (node2) -- (node6);

            \draw[->, btree-line] (node3) -- (node7);
            \draw[->, btree-line] (node3) -- (node8);

            \draw[->, btree-line] (node4) -- (node9);
            \draw[->, btree-line] (node4) -- (node10);

            \node[arrow-pointer, right=2cm of node9.center, rotate=-135, xshift=-2.7cm, yshift=-1cm] {};
            \node[draw, behind path, fit=(node4)(node9)(node10), rounded corners, dashed] {};
            \node[below=3cm of node3.center, minimum height=1cm] {};

            \draw[->, btree-line] ([yshift=-2cm]node4.center)
                .. controls ([yshift=-3cm] node4) and ([yshift=-3cm] node3) ..
                ([yshift=-2cm]node3.center);
            \draw[->, btree-line] ([yshift=-2cm]node4.center)
                .. controls ([yshift=-4cm] node4) and ([yshift=-4cm] node2) ..
                ([yshift=-2cm]node2.center);

        \end{tikzpicture}
        \linespread{0.5}
        \color{black}\fontsize{6pt}{0}\selectfont
            Riegger, Christian Vincon, Tobias Petrov, Ilia. (2017).
            Write-optimized indexing with partitioned b-trees. 296-300.
            10.1145/3151759.3151814.
        \linespread{1.5}

    \end{center}
\end{frame}
\note{
    For faster index creation, via pushing all merge runs into btree into
    different partitions.
}
